---
title: "Getting Started with RMySQL"
author: "Joseph Crispell"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    number_sections: true
    toc: yes
    toc_depth: '2'
    toc_float: yes
  html_notebook:
    highlight: haddock
    number_sections: yes
    theme: simplex
    toc: yes
    toc_depth: 2
    toc_float: yes
params: 
  password: "ThisIsNotMyPassword"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Usually a Reproducible Analytical Pipeline will start by retrieving data from a database. Often that database could be a `SQL` database. [`MySQL`](https://www.mysql.com/) is an open-source relational database management system that uses *S*tructured *Q*uery *L*anguage (SQL). We can use `MySQL` to store complex data tables that relate to one another.

`RMySQL` is an R package that we can use to interact with `MySQL`, writing queries, updating data and creating new tables.

The current tutorial aims to give a brief introduction to using `RMySQL` in R. The current content is more-or-less the same as the TutorialsPoint [tutorial](https://www.tutorialspoint.com/r/r_database.htm), with a  few updates and edits to match updates to `RMySQL`.

## Installing `MySQL`

To get started, you'll ned to install `MySQL`, here are links for the different operating systems:

- Installing on linux systems ([link](https://dev.mysql.com/doc/refman/8.0/en/linux-installation.html)), with extra help [here](https://itsfoss.com/install-mysql-ubuntu/).
- Installing on mac: ([link](https://dev.mysql.com/doc/mysql-osx-excerpt/5.7/en/osx-installation.html)), also, I think it comes ready installed! See [here](https://www.thoughtco.com/installing-mysql-on-mac-2693866))
- Installing on Windows ([link](https://dev.mysql.com/downloads/installer/)), with extra help [here](https://www.wikihow.com/Install-the-MySQL-Database-Server-on-Your-Windows-PC)

For this tutorial, I worked on a Windows machine and it took me a wee while to get MySQL installed properly, so be patient.

## Installing `RMySQL`

Much easier! Just use the following code:

```{r eval=FALSE}
install.packages("RMySQL")
```

Then we can load the library:
```{r}
library(RMySQL)
```


## Working with `RMySQL`

### Create a connection to MySQL

We are going to connect to the "sakila" database, which comes with the `MySQL` installation:

```{r eval=FALSE}
# Open a connection to a MySQL database
connection <- dbConnect(MySQL(), 
                            user='root', 
                            password=readline(prompt="Enter password: "), # Doing this as password never stored in easily accessible format now
                            dbname='sakila',
                            host = 'localhost')

# List the tables available in the sakila database
dbListTables(connection)
```


```{r echo=FALSE}
# Open a connection to a MySQL database
connection <- dbConnect(MySQL(), 
                            user='root', 
                            password=params$password, # Doing this as password never stored in easily accessible format now
                            dbname='sakila',
                            host = 'localhost')

# List the tables available in the sakila database
head(dbListTables(connection))
```

### Sending a query

We can query a table using the following code:

```{r}
# Query the "actor" table to get all the rows
result <- dbSendQuery(conn=connection, statement=SELECT * FROM actor") # * means everything NOTE: syntax is case-insensitive

# Store the first 15 rows of the result in an R data frame object
actors <- dbFetch(result, n=15)

# Remove the unused results
dbClearResult(result)
```

Note that the last line of code (`dbClearResult(result)`) clears the fecthed data and closes the query. It is now preferred to query a table with the following code, which doesn't leave the query open:

```{r}
# Note - you can use single command below to send query and send result as data.frame in one command - doesn't leave query open
actors <- dbGetQuery(conn=mySQLConnection, statement="SELECT * FROM actor WHERE last_name = 'TEMPLE'")

```

