---
title: "Getting Started with RMySQL"
author: "Joseph Crispell"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    number_sections: true
    toc: yes
    toc_depth: '2'
    toc_float: yes
params: 
  password: "ThisIsNotMyPassword"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Usually a Reproducible Analytical Pipeline will start by retrieving data from a database. Often that database could be a `SQL` database. [`MySQL`](https://www.mysql.com/) is an open-source relational database management system that uses *S*tructured *Q*uery *L*anguage (SQL). We can use `MySQL` to store complex data tables that relate to one another.

`RMySQL` is an R package that we can use to interact with `MySQL`, writing queries, updating data and creating new tables.

The current tutorial aims to give a brief introduction to using `RMySQL` in R. The current content is more-or-less the same as the TutorialsPoint [tutorial](https://www.tutorialspoint.com/r/r_database.htm), with a  few updates and edits to match updates to `RMySQL`.

# Installing `MySQL`

To get started, you'll need to install `MySQL`, here are links for the different operating systems:

- Installing on linux systems ([link](https://dev.mysql.com/doc/refman/8.0/en/linux-installation.html)), with extra help [here](https://itsfoss.com/install-mysql-ubuntu/).
- Installing on mac: ([link](https://dev.mysql.com/doc/mysql-osx-excerpt/5.7/en/osx-installation.html)), also, I think it comes ready installed! See [here](https://www.thoughtco.com/installing-mysql-on-mac-2693866))
- Installing on Windows ([link](https://dev.mysql.com/downloads/installer/)), with extra help [here](https://www.wikihow.com/Install-the-MySQL-Database-Server-on-Your-Windows-PC)

For this tutorial, I worked on a Windows machine and it took me a wee while to get MySQL installed properly, so be patient. Also, I had to use the legacy password based authentification to get R to connect properly (see this [question](https://stackoverflow.com/questions/49931541/mysql-changing-authentication-type-from-standard-to-caching-sha2-password) for more information).

# Installing `RMySQL`

Much easier! Just use the following code:

```{r eval=FALSE}
install.packages("RMySQL")
```

Then we can load the library:
```{r}
library(RMySQL)
```


# Working with `RMySQL`

## Create a connection to MySQL

We are going to connect to the "sakila" database, which comes with the `MySQL` installation:

```{r eval=FALSE}
# Open a connection to a MySQL database
connection <- dbConnect(MySQL(),
  user = "root",
  password = readline(prompt = "Enter password: "), # Password not in code
  dbname = "sakila",
  host = "localhost"
)

# List the tables available in the sakila database
dbListTables(connection)
```


```{r echo=FALSE}
# Open a connection to a MySQL database
connection <- dbConnect(MySQL(),
  user = "root",
  password = params$password, # Password not stored in code
  dbname = "sakila",
  host = "localhost"
)

# List the tables available in the sakila database
head(dbListTables(connection))
```

## Sending a query

We can query a table using the following code:

```{r warning=FALSE}
# Query the "actor" table to get all the rows
# NOTE: SQL syntax is case-insensitive
result <- dbSendQuery(conn = connection, statement = "SELECT * FROM actor")

# Store the first 15 rows of the result in an R data frame object
actors <- dbFetch(result, n = 15)

# Remove the unused results
dbClearResult(result)
```

Note that the last line of code (`dbClearResult(result)`) clears the fetched data and closes the query. It is now preferred to query a table with the following code, which doesn't leave the query open:

```{r warning=FALSE}
# Note - you can use single command below to send query and send result
# as data.frame in one command - doesn't leave query open
actors <- dbGetQuery(conn = connection, statement = "SELECT * FROM actor")
```

Let's take a quick look at the actors table:
```{r}
head(actors)
```


Now, let's try getting the information for actors whose last name is "TEMPLE":
```{r warning=FALSE}
# Note - you can use single command below to send query and send result
# as data.frame in one command - doesn't leave query open
actors <- dbGetQuery(
  conn = connection,
  statement = "SELECT * FROM actor WHERE last_name = 'TEMPLE'"
)

# Print the table
(actors)
```

## Creating a table

In order to create a `MySQL` table, we first need to set our permissions so that we can input data locally (see [this](https://stackoverflow.com/questions/44288358/is-there-a-faster-way-to-upload-data-from-r-to-mysql) question for more information):

```{r}
dbSendQuery(conn = connection, statement = "SET GLOBAL local_infile = 1")
```

With that done, we are going to store the `mtcars` table on our local `MySQL` server using the following code:
```{r}
# Upload mtcars table onto MySQL server
dbWriteTable(connection, "mtcars", mtcars[, ], overwrite = TRUE)

# Check the table is there
mtcars_mysql <- dbGetQuery(
  conn = connection,
  statement = "SELECT * FROM mtcars"
)
head(mtcars_mysql)
```


## Updating a table

We can update a table on our `MySQL` server by sending a query, for example we can use the code below to edit the data for cars with a certain horse power:
```{r}
# Update data in the actor info table
dbSendQuery(
  conn = connection,
  statement = "UPDATE mtcars SET disp = 168.5 WHERE hp = 110"
)

# Check the updated table
mtcars_mysql <- dbGetQuery(
  conn = connection,
  statement = "SELECT * FROM mtcars"
)

# Take a look to see if the data has changed
head(mtcars_mysql[mtcars_mysql$hp == 110, ])
```

## Inserting a row into a table

Using the code below, we can insert data for a new car into the `mtcars` table on our `MySQL` server:
```{r}
# Insert a new row into table
dbSendQuery(
  conn = connection,
  statement = paste0(
    "INSERT INTO mtcars(row_names, mpg, cyl, disp, hp, drat,",
    "wt, qsec, vs, am, gear, carb) values('New Mazda RX4 Wag',",
    "21, 6, 168.5, 110, 3.9, 2.875, 17.02, 0, 1, 4, 4)"
  )
)

# Check the updated table
mtcars <- dbGetQuery(conn = connection, statement = "SELECT * FROM mtcars")
tail(mtcars)
```

## Removing a table

We can remove the `mtcars` table from our `MySQL` server with the following code:
```{r}
dbSendQuery(connection, "drop table if exists mtcars")
```

## Closing a connection to the `MySQL` server

To wrap up, we can close our connection to the `MySQL` server using the following code:
```{r warning=FALSE}
dbDisconnect(conn = connection)
```

# The end :-)

So, that's me done. This tutorial has hopefully got you started with using the `RMySQL` R package. If you haven't already I would recommend going and doing some more tutorials about `MySQL`. [Here](https://www.tutorialspoint.com/mysql/index.htm) is a good one to start with.

As you'll have noticed, this tutorial has barely scratched the surface of using `MySQL`. It is an extremely powerful language and one that will be found in many Reproducible Analytical Pipelines and well worth learning about.
