---
title: "Documenting skills development"
author: "Joseph Crispell & Laurie Baker"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: '2'
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preparation, include=FALSE}
# Load libraries
library(fmsb)
```

# Introduction

Here is a quick tutorial about mapping our skills progression using R. We're going to be using the [fmsb]() to visualise our skills progress, which we'll install and load with the following code:

```{r install and load, eval=FALSE}
install.packages("fmsb")
library(fmsb)
```

# Our skill areas

We'll look to categorise our skills progress into the following themes:

- **Programming**
- **Statistics**
- **Databases**
- **Projects**
- **Web development**
- **Versioning**

We can then create a `data.frame` that records our current skills point:

```{r define skills table}
# Set you current skills point
skillsPointsCurrent <- data.frame("Programming"=c(5, 1, 4), 
                                  "Statistics"=c(5, 1, 3),
                                  "Databases"=c(5, 1, 2),
                                  "Projects"=c(5, 1, 3),
                                  "Web"=c(5, 1, 3),
                                  "Versioning"=c(5, 1, 3))
rownames(skillsPointsCurrent) <- c("Max level", "Min level", "My level")

skillsPointsCurrent
```

# Visualising our skills point

```{r}
radarchart(skillsPointsCurrent, 
           axistype=1, # Centre axis labels
           title="Starting Point",
           pcol=rgb(0.2,0.5,0.5, 0.6), # Colour of polygon points
           pfcol=rgb(0.2,0.5,0.5, 0.5), # COlour of polygon fill
           plwd=2, # Line width for the polygon
           cglcol="grey", # COlour of lines in web
           cglty=1, # Type of line in radar
           axislabcol="blue", # Skills categories colour
           caxislabels=c("Unaware","Aware","Working","Practitioner", "Expert"), # Skills categories
           cglwd=0.3, # Width of line for radar
           calcex=0.75, # Font size for axis labels
           vlcex=1) # Font size for skills labels
```


# Thinking about progression

Next, I'd like to consider where I would like to get to with my learning:

```{r where Id like to be}
# Set you goal skills point
skillsPointsGoal <- data.frame("Programming"=c(5, 1, 4), 
                               "Statistics"=c(5, 1, 4),
                               "Databases"=c(5, 1, 3),
                               "Projects"=c(5, 1, 4),
                               "Web"=c(5, 1, 3),
                               "Versioning"=c(5, 1, 5))
rownames(skillsPointsGoal) <- c("Max level", "Min level", "My level")

# Plot my goal point
radarchart(skillsPointsGoal, 
           axistype=1, # Centre axis labels
           title="Where I'd like to be",
           pcol=rgb(0.2,0.5,0.5, 0.6), # Colour of polygon
           pfcol=rgb(0.2,0.5,0.5, 0.5), # COlour of polygon fill
           plwd=2, # Line width for the polygon
           cglcol="grey", # COlour of lines in web
           cglty=1, # Type of line in radar
           axislabcol="blue", # Skills categories colour
           caxislabels=c("Unaware","Aware","Working","Practitioner", "Expert"), # Skills categories
           cglwd=0.3, # Width of line for radar
           calcex=0.75, # Font size for axis labels
           vlcex=1) # Font size for skills labels
```

# Overlaying skills points in a single plot

We can use the [basicPlotteR](https://github.com/JosephCrispell/basicPlotteR/blob/master/README.md#radarchart) R package (shameless self-promotion!) to overlay our skills points in a single plot. Install it with the following code:

```{r install basicPlotteR, eval=FALSE}
install.packages("devtools")
devtools::install_github("JosephCrispell/basicPlotteR")
library(basicPlotteR)
```

```{r load basicPlotteR, echo=FALSE}
library(basicPlotteR)
```

Then create the radar chart with the following code:

```{r radar chart with basicPlotteR}
# Create a chart with my current skills points
radarChart(scores=skillsPointsCurrent["My level", ],
           names=colnames(skillsPointsCurrent), 
           levels=c("Unaware","Aware","Working","Practitioner", "Expert"))

# Add my goal onto the same chart
radarChart(scores=skillsPointsGoal["My level", ],
           names=colnames(skillsPointsGoal), 
           levels=c("Unaware","Aware","Working","Practitioner", "Expert"), add=TRUE,
           polygon.col="blue")

# Add a legend
legend("topright", legend=c("Current", "Goal"), text.col=c("red", "blue"),  bty="n", xpd=TRUE)
```

# Creating an interactive skills map

Lastly, we can use the [plotly]() R package to create an interactive radar chart. Install plotly with the following code:

```{r install plotly, eval=FALSE}
install.packages("plotly")
library(plotly)
```

```{r load plotly, include=FALSE}
library(plotly)
```

Then create your interactive chart like this:

```{r radar chart with plotly, warning=FALSE}
# Create a radial plot object
fig <- plot_ly(type="scatterpolar", fill="toself", mode="markers") 

# Add the first skills point trace - current
fig <- add_trace(fig, 
                 r=as.numeric(skillsPointsCurrent["My level", ]),
                 theta=colnames(skillsPointsCurrent),
                 name="Current") 

# Add the second skills point trace - my goals
fig <- add_trace(fig,
                 r=as.numeric(skillsPointsGoal["My level", ]),
                 theta=colnames(skillsPointsGoal),
                 name="Goal")

# Add a radial axis
fig <- layout(fig,
              polar=list(
                radialaxis=list(visible=TRUE,
                                range=c(1,5),
                                tickvals=1:5,
                                ticktext=c("Unaware","Aware","Working","Practitioner", "Expert"))
                )
              )

# Plot the radial chart
fig
```



NEED TO LOOK INTO DOING THIS PLOTLY: https://plotly.com/r/radar-chart/