---
title: "Reproducible Analytical Pipeline"
author: "Joseph Crispell & Nathan Begbie"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  ioslides_presentation:
    logo: Images/DataScienceCampus_logo_square.png
    widescreen: true
  beamer_presentation: default
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<!-- Extra custom code to add "OFFICIAL" into the header -->

<script src="https://ajax.googleapis.com/ajax/libs/jquery/1.12.2/jquery.min.js"></script>

<script>
    $(document).ready(function() {
      $('slide:not(.title-slide, .backdrop, .segue)').append('<header label=\"OFFICIAL\"></header>');    
    })
</script>

<style>
  header:after {
    content: attr(label);
    font-size: 12pt;
    position: absolute;
    top: 10px;
    left: 10px;
    line-height: 1.9;
  }
</style>

<!-- END custom code block for header -->


## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

