---
title: "RShiny and flex_dashboard example"
author: "Joseph Crispell"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
    flexdashboard::flex_dashboard:
      orientation: columns
      source_code: embed
runtime: shiny
---

```{r global, include=FALSE}

# Using 'global' chunk, so that all data can be loaded can be shared by all users of the dashboard
library(datasets) # Load datasets package - example datasets
data(faithful) # Load eruption data
```

Guide
=====================================

### Introduction

The current document provides some simple examples that combine `flex_dashboard` and `RShiny` to create interactive dashboards for visualising and exploring data. As you can see this document is a work in progress!!

### Useful resources

Here's a list of resources that I found really useful when I was creating this dashboard:

- [Using `RShiny` and `flex_dashboard`](https://rmarkdown.rstudio.com/flexdashboard/shiny.html)
- [Getting started with `flex_dashboard`](https://rmarkdown.rstudio.com/flexdashboard/using.html)
- [Examples of `RShiny` and `flex_dashboard` in action](https://rmarkdown.rstudio.com/flexdashboard/examples.html)
- [Creating an `RShiny` `flex_dashboard`](https://medium.com/analytics-vidhya/shiny-dashboards-with-flexdashboard-e66aaafac1f2)


Example: Geyser
=====================================

Column {.sidebar}
-----------------------------------------------------------------------

Waiting time between eruptions and the duration of the eruption for the
Old Faithful geyser in Yellowstone National Park, Wyoming, USA.

```{r}
selectInput("n_breaks", label = "Number of bins:",
            choices = c(10, 20, 35, 50), selected = 20)

sliderInput("bw_adjust", label = "Bandwidth adjustment:",
            min = 0.2, max = 2, value = 1, step = 0.2)
```

Column
-----------------------------------------------------------------------

### Geyser Eruption Duration

```{r}
renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser Eruption Duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```
