---
title: "Documenting skills development"
author: "Joseph Crispell & Laurie Baker"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: '2'
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preparation, include=FALSE}
# Load libraries
library(fmsb)
```

# Introduction

Here is a quick tutorial about mapping our skills progression using R. We're going to be using the [fmsb]() to visualise our skills progress, which we'll install and load with the following code:

```{r install and load, eval=FALSE}
install.packages("fmsb")
library(fmsb)
```

# Our skill areas

We'll look to categorise our skills progress into the following themes:

- **Programming**
- **Statistics**
- **Databases**
- **Projects**
- **Web development**
- **Versioning**

We can then create a `data.frame` that records our current skills point:

```{r define skills table}
# Set you current skills point
skillsPointsCurrent <- data.frame("Programming"=c(5, 1, 4), 
                                  "Statistics"=c(5, 1, 3),
                                  "Databases"=c(5, 1, 2),
                                  "Projects"=c(5, 1, 3),
                                  "Web"=c(5, 1, 3),
                                  "Versioning"=c(5, 1, 3))
rownames(skillsPointsCurrent) <- c("Max level", "Min level", "My level")

skillsPointsCurrent
```

# Visualising our skills point

```{r}
radarchart(skillsPointsCurrent, 
           axistype=1, # Centre axis labels
           title="Starting Point",
           pcol=rgb(0.2,0.5,0.5, 0.6), # Colour of polygon points
           pfcol=rgb(0.2,0.5,0.5, 0.5), # COlour of polygon fill
           plwd=2, # Line width for the polygon
           cglcol="grey", # COlour of lines in web
           cglty=1, # Type of line in radar
           axislabcol="blue", # Skills categories colour
           caxislabels=c("Unaware","Aware","Working","Practitioner", "Expert"), # Skills categories
           cglwd=0.3, # Width of line for radar
           calcex=0.75, # Font size for axis labels
           vlcex=1) # Font size for skills labels
```


# Thinking about progression

Next, I'd like to consider where I would like to get to with my learning:

```{r where Id like to be, echo=FALSE}
# Set you goal skills point
skillsPointsGoal <- data.frame("Programming"=c(5, 1, 4), 
                               "Statistics"=c(5, 1, 4),
                               "Databases"=c(5, 1, 3),
                               "Projects"=c(5, 1, 4),
                               "Web"=c(5, 1, 3),
                               "Versioning"=c(5, 1, 5))
rownames(skillsPointsCurrent) <- c("Max level", "Min level", "My level")

# Plot my goal point
radarchart(skillsPointsGoal, 
           axistype=1, # Centre axis labels
           title="Where I'd like to be",
           pcol=rgb(0.2,0.5,0.5, 0.6), # Colour of polygon
           pfcol=rgb(0.2,0.5,0.5, 0.5), # COlour of polygon fill
           plwd=2, # Line width for the polygon
           cglcol="grey", # COlour of lines in web
           cglty=1, # Type of line in radar
           axislabcol="blue", # Skills categories colour
           caxislabels=c("Unaware","Aware","Working","Practitioner", "Expert"), # Skills categories
           cglwd=0.3, # Width of line for radar
           calcex=0.75, # Font size for axis labels
           vlcex=1) # Font size for skills labels
```

NEED TO LOOK INTO DOING THIS PLOTLY: https://plotly.com/r/radar-chart/