---
title: "A stand-a-lone interactive Rmarkdown dashboard"
author: "Joseph Crispell"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
---

```{r setup, include=FALSE}
# Load the required libraries
library(flexdashboard) # Create Rmarkdown flexdashboard
library(plotly) # Interative plots
library(kableExtra) # Extra nice tables
library(RColorBrewer) # Creating colour pallettes
library(DT) # An interactive table format

# Load the common functions
source("ExampleDashboards_FUNCTIONS.R")
```


Main
=======================================================================

Column {data-width=350}
-----------------------------------------------------------------------

### Getting started

The current document aims to illustrate how we can used [Rmarkdown](https://rmarkdown.rstudio.com/) to create interactive documents for visualising and exploring data. We'll be exploring the [COVID-19](https://covid19.who.int/) data published by the World Health Organisation. The World Health Organisation publishes COVID-19 case and death counts for most countries. We can load these data with the following code:

```{r load the data, echo=TRUE}
# Note the url for the WHO COVID-19 data
whoURL <- "https://covid19.who.int/WHO-COVID-19-global-data.csv"

# Load the WHO COVID-19 data
covid <- read.csv(whoURL)

# Fix the first column name
colnames(covid)[1] <- "Date_reported"

# Convert the dates from character vector to a date object
covid$Date_reported <- as.Date(covid$Date_reported, format="%Y-%m-%d")
```

The WHO COVID-19 data looks like this:
```{r view WHO data}
# Take a quick look at the table - note I am limiting view to first 25 rows
prettyTable(covid[1:25, ])
```

### Processing the data

The WHO data reports the new and cumulative counts for cases and deaths for each country since the start of January. We want to visualise the daily trends in deaths for each country, so we'll need to manipulate the data a little to align the daily deaths counts by date (*note that I am using a bespoke function to do this processing*):
```{r align by date, echo=TRUE}
# Align the daily death counts by date
alignedByDate <- alignCountByDate(covid, countColumn="New_deaths")
```

The aligned data looks like this:
```{r view aligned data}
# Take a quick look at the aligned deaths data - note only selecting 25 rows and 25 countries
prettyTable(alignedByDate[100:125, 1:25])
```

Column {data-width=650}
-----------------------------------------------------------------------

### Visualising trends  countries with highest number of total deaths

```{r plot the daily deaths}
# Select the countries with the highest cumulative death counts
nToSelect <- 5
topCountries <- getTopCountries(covid, countColumn="Cumulative_deaths", n=nToSelect)

# Ploit the trends for the countries with highest counts
plotDailyDeathsInteractive(aligendByDate, countries=topCountries, 
                           title=paste0("Trends in daily deaths in top ", nToSelect, " countries with highest total deaths"))
```

Data
=======================================================================

```{r interactive table}
datatable(alignedByDate[, c("Date", topCountries)], rownames=FALSE)
```

United Kingdom {data-navmenu="Countries"}
=======================================================================

```{r, echo=FALSE}
backgroundCountries <- colnames(alignedByDate)[-1]
backgroundCountries <- backgroundCountries[backgroundCountries != "The United Kingdom"]
plotDailyDeathsInteractive(aligendByDate, countries=c("The United Kingdom"),
                           countriesInBackground=backgroundCountries,
                           title=paste0("Trends in daily deaths in UK"),
                           legend=FALSE)
```

United States of America {data-navmenu="Countries"}
=======================================================================

```{r, echo=FALSE}
backgroundCountries <- colnames(alignedByDate)[-1]
backgroundCountries <- backgroundCountries[backgroundCountries != "United States of America"]
plotDailyDeathsInteractive(aligendByDate, countries=c("United States of America"),
                           countriesInBackground=backgroundCountries,
                           title=paste0("Trends in daily deaths in US"),
                           legend=FALSE)
```

China {data-navmenu="Countries"}
=======================================================================

```{r, echo=FALSE}
backgroundCountries <- colnames(alignedByDate)[-1]
backgroundCountries <- backgroundCountries[backgroundCountries != "China"]
plotDailyDeathsInteractive(aligendByDate, countries=c("China"),
                           countriesInBackground=backgroundCountries,
                           title=paste0("Trends in daily deaths in China"),
                           legend=FALSE)
```

Resources
=======================================================================

Here's some resources that I found useful when creating this document:

- [Getting started with Rmarkdown](https://rmarkdown.rstudio.com/lesson-1.html)
- [Creating a dashboard with `flexdashboard`](https://rmarkdown.rstudio.com/flexdashboard/using.html)
- [Rmarkdown cheatsheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
- [Interactive plotting with plotly](https://plotly.com/r/)
- [Programming in R](https://www.tutorialspoint.com/r/index.htm)
- [Troubleshooting with stackoverflow](https://stackoverflow.com/questions/tagged/r)


