---
title: "Testing connection to MySQL"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    runtime: shiny
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
# Load libraries
library(flexdashboard) # Building the dashboard
library(shiny) # Adding interactive elements for user input
library(RMySQL) # Interacting with local MySQL server
library(DBI) # Used by RMySQL

# Get MySQL password
sqlPassword <- Sys.getenv("MYSQLPASS") # I have stored as environmental variable
sqlUser <- "root"
```

```{r define functions to interact with MySQL server}

connect <- function(password, user = "root"){
  
  # Connect to MySQL server
  connection <- dbConnect(
      drv = MySQL(),
      user = user, 
      password = password,
      dbname = 'sakila',
      host = 'localhost')
  
  return(connection)
}

disconnect <- function(connection){
  dbDisconnect(connection)
}

sendMySQLQuery <- function(query, password, user = "root"){
  
  # Connecting/disconnecting for each query because shiny was breaking after 16 connections made (see question here: https://stackoverflow.com/questions/33885400/rmysql-internal-error-in-rs-dbi-getconnection-corrupt-connection-handle)
  
  # Connect to MySQL server
  connection <- connect(password, user)
  
  # Send query to MySQL database
  result <- dbGetQuery(connection, query)
  
  # Disconnect from MySQL server
  disconnect(connection)
  
  return(result)
}
```


Trade Trends by Countries
========================================

Inputs {.sidebar}
----------------------------------------

```{r user input}

numericInput("nrows", "Enter the number of rows to display:", 5)
selectInput("actor", "select actor", c("NICK", "ED"))
```

Column
------------------------------------

Below should a simple table giving details of actors (taken from the `actor` table in the [`sakila`](https://dev.mysql.com/doc/sakila/en/) database that comes with every MySQL download). The number of rows in the table should be set to no more than ``r reactive({input$nrows})`` (set by user on the left). The actor's detailed in the table below should all have the first_name of ``r reactive({input$actor})`` (set by user on the left).

```{r get data based on user input}

# Select data from MySQL server
actorInfo <- reactive({

  sendMySQLQuery(paste0("SELECT * FROM actor WHERE first_name = \"", input$actor, "\" LIMIT ", input$nrows, ";"), password = sqlPassword, user = sqlUser)
})
```

```{r show data}
# View the actor data
renderTable({
  actorInfo()
})
```
