---
title: "Monitoring KNBS national census sampling frame (last updated: `r format(Sys.Date(), '%d %b %Y')`)"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
---

```{r preparation}
# Load required libraries
library(leaflet) # Interactive maps
library(plotly) # Interactive plots
library(flexdashboard) # Value boxes
```

Sidebar {.sidebar}
=======================================================================

### An example flexdashboard

The current dashboard is a demonstration of the type of dashboard that can be built using [Rmarkdown](https://rmarkdown.rstudio.com/lesson-1.html) and [flexdashboard](https://pkgs.rstudio.com/flexdashboard/). Rmarkdown is the underlying script - a combination of R and [markdown](https://opensource.com/article/19/9/introduction-markdown) to enable effective reporting. Flexdashboard is a library that makes setting up the structure of your dashboard a little easier.


Dashboard
=======================================================================

Row
-----------------------------------------------------------------------

### Number of clusters visited {.value-box}

```{r number clusters}
valueBox(
  value = 250,
  icon = "fa-check-circle"
)
```

### Number of surveyors on the ground {.value-box}

```{r number surveyors}
valueBox(
  value = 25,
  icon = "fa-shoe-prints"
)
```

### Current phase {.value-box}

```{r current phase}
valueBox(
  value = 2,
  caption = "Current phase",
  icon = "fa-calendar-check"
)
```

Row
-----------------------------------------------------------------------

### An example of interactive map

```{r leaflet map}
# Example taken from: http://rstudio.github.io/leaflet/

# Initialise the leaflet map
interactive_map <- leaflet() %>%
  # Add background map
  addTiles() %>%
  # Add point location
  addMarkers(lng = 174.768, lat = -36.852, popup = "The birthplace of R")

# Show map
interactive_map
```

### An example of an interactive plot

```{r plot plot}
# Generate some example time series
trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)

# Generate X axis values
x <- c(1:100)

# Store example data in data frame
data <- data.frame(x, trace_0, trace_1, trace_2)

# Initialise plotly figure with first time series
fig <- plot_ly(data,
  x = ~x, y = ~trace_0,
  name = "trace 0", type = "scatter",
  mode = "lines"
)

# Add additional traces (lines)
fig <- fig %>% add_trace(y = ~trace_1, name = "trace 1", mode = "lines+markers")
fig <- fig %>% add_trace(y = ~trace_2, name = "trace 2", mode = "markers")

# Show figure
fig
```

Page 2
=======================================================================
