---
title: "Creating a GANTT chart in R"
author: "Joseph Crispell"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    number_sections: true
    toc: yes
    toc_depth: '2'
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r set-options, echo=FALSE, cache=FALSE}
options(width=800)
```

```{r preparation, include=FALSE}
# Load libraries
library(knitr) # Nice tables
library(kableExtra) # Extra nice tables
library(timevis) # Wrapper for javascript gantt library: https://visjs.org/
library(htmlwidgets) # Creating html widgets
library(tufte) # Block quotes

# Set the path
path <- file.path("C:", "Users", "j-crispell", "Repositories", "gov-uk-rap-materials", "CreatingAGantt")
```

```{r functions, include=FALSE}
prettyTable <- function(table){
  
  # Use the kable function to create a nicer formatted table
  kable(table) %>%
  
    # Set the format
    kable_styling(bootstrap_options="striped", # Set the colour of rows
                  full_width=FALSE, # Make the table not stretch to fit the page
                  position="left") %>% # Position the table on the left
  
    # Make the table scrollable
    scroll_box(height = "400px")
}
```

```{r load gantt, echo=FALSE}
# Read in the GANTT table
file <- file.path(path, "GanttTutorialTimeline_18-05-20.csv")
gantt <- read.table(file, header=TRUE, sep=",", stringsAsFactors=FALSE)

# Set the date formats
gantt$start <- as.POSIXct(gantt$start, format="%d/%m/%Y %H:%M")
gantt$end <- as.POSIXct(gantt$end, format="%d/%m/%Y %H:%M")

# Remove NA rows
naRows <- which(is.na(gantt$id))
if(length(naRows) > 0){
  gantt <- gantt[-naRows, ]
}
```

```{r create gantt html, echo=FALSE}
# Note the groupings in the gannt chart
groups <- c("Preparation", "Presenting")
style <- c("<b>Preparation</b>", "<b>Presenting</b>")
groups <- data.frame(id=groups, content=style)

# Set the options for timevis
options <- list(editable=FALSE, align="center", multiselect=TRUE)

# Save the timeline to file
file <- file.path(path, "GanttTutorial_gantt.html")
saveWidget(timevis(gantt, groups=groups, options=options, width="100%", showZoom=TRUE), 
           file, selfcontained=FALSE) # Unfortunately self-contained doesn't work
```

# What is a gantt chart

It's A horizontal bar chart illustrating a project schedule. It was first used by Henry Gantt in ~1910.

Here is a gantt chart mapping how I will create the current tutorial:

```{r show gantt, echo=FALSE, fig.width=10}

# Plot the GANTT chart
timevis(gantt, groups=groups, options=options)
```

The above gantt chart sets out the work packages (`WP`) and their associated tasks (`T`) and deliverables (`D`).

# Creating a gantt chart in R

The current tutorial will describe how to create an interactive html-based gantt chart in R. We'll be using the `timevis` R package (really nice demo [here](https://daattali.com/shiny/timevis-demo/)). `timevis` is a wrapper for the `vis.js` javascript visulisation library (described [here](https://visjs.org/)).

## Create the timeline table

Any gantt chart needs a table that sets out the timings and duration of each element in the chart:

```{r show input table, echo=FALSE}
prettyTable(gantt[, c("content", "start", "end")])
```

<br>

Work packages (`WP`s) and tasks (`T`s) have both a start and end date/time and take a period of time. Deliverables (`D`s) have a start date - the due date for the output.

For my gantt chart, I have included a few more columns to set some style and groupings that `timevis` can use:

```{r show full input table, echo=FALSE, fig.width=10}
prettyTable(gantt)
```

<br>

Here is a brief description of each of the columns:

- **id**: a unique identifying number for each element
- **content**: text to be shown in gantt
- **title**: text to be shown in tooltip when hovering
- **start**: date/time element is due to start
- **end**: date/time element is due to end
- **style** - `css` code used to set style of element on gantt
- **group** - categorical column dividing elements into groups

I created the above table in excel and saved as `csv` formatted text file to read into R.

## Read in timeline table into R and process

We can read in the timeline table with the following code:

```{r load timeline table, eval=FALSE, echo=TRUE}
# Read in the GANTT table
file <- file.path(path, "GanttTutorialTimeline_18-05-20.csv")
gantt <- read.table(file, header=TRUE, sep=",", stringsAsFactors=FALSE)
```

<br>
Next, we can note the date columns and remove `NA` rows (excel kept inserting these for me!).

```{r note date columns and remove NA rows, eval=FALSE, echo=TRUE}
# Set the date formats
gantt$start <- as.POSIXct(gantt$start, format="%d/%m/%Y %H:%M")
gantt$end <- as.POSIXct(gantt$end, format="%d/%m/%Y %H:%M")

# Remove NA rows
naRows <- which(is.na(gantt$id))
if(length(naRows) > 0){
  gantt <- gantt[-naRows, ]
}
```

<br>

We can recognise the grouping column and use some `html` to set some styles for their labels:

```{r note the group column, eval=FALSE, echo=TRUE}
# Note the groupings in the gannt chart
groups <- c("Preparation", "Presenting")
style <- c("<b>Preparation</b>", "<b>Presenting</b>")
groups <- data.frame(id=groups, content=style)
```

## Plotting the gantt

Now we can visualise our gantt chart in R with the following code:

```{r visualise gantt, eval=TRUE, echo=TRUE, fig.width=10}
# Set the options for timevis
options <- list(editable=FALSE, align="center", multiselect=TRUE)

# Plot the GANTT chart
timevis(gantt, groups=groups, options=options)
```

We can also save the gantt as an `html` file:
```{r save gantt as html, eval=FALSE, echo=TRUE}
# Save the timeline to file
file <- file.path(path, "GanttTutorial_gantt.html")
saveWidget(timevis(gantt, groups=groups, options=options, 
                   width="100%", showZoom=TRUE), 
           file, selfcontained=FALSE) # Unfortunately self-contained doesn't work well for me
```

# Wrapping up

So `timevis` is a cool R package to help us create interactive gantt charts in R. Gantt charts are useful for planning projects and monitoring their progress.

The R code used in the current tutorial for creating a gantt is available [here](https://github.com/datasciencecampus/gov-uk-rap-materials/blob/master/CreatingAGantt/CreateGanttChart_18-05-20.R).

Hope this is useful!
